PLAE enables quick and easy searching across one million ocular transcriptomes

### 2022-06-24 manuscript_00.Rmd
more

# number of celltype (predict) - organism - study counts (with >=50 cells)
ws_table <- meta_filter %>%  
  filter(Organ == 'Eye') %>% 
  group_by(organism, CellType_predict, study_accession) %>% 
  count() %>% 
  filter(n >= 50) %>% 
  ungroup() %>% 
  group_by(organism, CellType_predict) %>% 
  count() %>% 
  pivot_wider(values_from = n, names_from = organism) %>% 
  filter(!is.na(CellType_predict)) %>% 
  replace(is.na(.), values = 0) %>% 
  arrange(CellType_predict) %>% 
  rename(`CellType (Predict)` = CellType_predict) %>% 
  mutate(`Well Supported` = case_when(`Homo sapiens` >= 2 & `Mus musculus` >= 2 ~ 'Yes',
                                      TRUE ~ 'No'))
ws_table %>% 
  flextable()
```

@@

# id diff genes only seen once across the cell types
uniq_g <- consist_diff %>% group_by(Symbol) %>% count() %>% filter(n == 1)
consist_diff %>% group_by(Base) %>% summarise(Count = n())
consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n())
```

@@

hm <- ComplexHeatmap::Heatmap(x, column_labels = organism, 
                              column_split = cts, 
                              row_split = base_yo, show_row_dend = FALSE, 
                              row_labels = genes, 
                              column_title_rot = 90, 
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
ComplexHeatmap::draw(hm, padding = unit(c(2, 2, 30, 2), "mm"))
```

@@

Neurogenic and RPC have very similar expression patterns. We leverage our more specific differential test to directly assess genes were are different between neurogenic cells and RPC. We show the top 5 genes that are both more highly expressed in RPC relative to neurogenic cell and vice versa.
```{r}
x <-  scEiaD_2020_v01 %>% 
    tbl("diff_testing") %>% 
    filter(Against  == "RPC",   
           Group == 'CellType (Predict)', 
           Base == "Neurogenic Cell") %>% 
  group_by(Base, Against, Gene) %>% 
  summarise(`mean log2FC` = mean(log2FoldChange), `mean padj` = mean(padj), `sig count` = sum(pvalue < 1e-2), `na count` = sum(is.na(log2FoldChange))) %>% 
  arrange(-`mean log2FC`) %>%   
  collect() %>% 
  filter(!is.na(`mean log2FC`))

@@

```{r}
# consistently diff genes
consistent <- diff_tab %>% 
  group_by(Symbol, Base)  %>% 
  summarise(Count = n()) %>% 
  filter(Count > 1) %>% 
  mutate(k = paste(Symbol, Base, sep = '_'))

diff_tab %>% mutate(k = paste(Symbol, Base, sep = '_')) %>% 
  filter(k %in% consistent$k) %>% 
  group_by(Base, Organism) %>% 
  summarise(Count = n()) %>% 
  rename(`CellType (Predict)` = Base, `Differentially Expressed Gene Count` = Count) %>% 
  pivot_wider(values_from = `Differentially Expressed Gene Count`, names_from = Organism) %>% replace(is.na(.),0) %>% 
  as_tibble() %>% 
  flextable()

@@

hm <- ComplexHeatmap::Heatmap(x, column_labels = organism, 
                              row_split = base_yo, show_row_dend = FALSE, 
                              row_labels = genes, 
                              column_title_rot = 90, 
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
hm
```

>>>>

# number of celltype (predict) - organism - study counts (with >=50 cells)
ws_table <- meta_filter %>%  
  filter(Organ == 'Eye') %>% 
  group_by(organism, CellType_predict, study_accession) %>% 
  count() %>% 
  filter(n >= 50) %>% 
  ungroup() %>% 
  group_by(organism, CellType_predict) %>% 
  count() %>% 
  pivot_wider(values_from = n, names_from = organism) %>% 
  filter(!is.na(CellType_predict)) %>% 
  replace(is.na(.), values = 0) %>% 
  arrange(CellType_predict) %>% 
  rename(`CellType (Predict)` = CellType_predict) %>% 
  mutate(`Well Supported` = case_when(`Homo sapiens` >= 2 & `Mus musculus` >= 2 ~ 'Yes',
                                      TRUE ~ 'No'))
ws_table %>% 
  flextable()

diff_tab_sup %>% pivot_longer(-`CellType (Predict)`, names_to = 'Organism', values_to = 'Diff Gene Count') %>% ggplot(aes(x=`CellType (Predict)`, y = `Diff Gene Count`, fill = Organism )) + geom_bar(stat='identity', position = 'dodge') + cowplot::theme_cowplot() + coord_flip()
```

@@

# id diff genes only seen once across the cell types
uniq_g <- consist_diff %>% group_by(Symbol) %>% count() %>% filter(n == 1)
consist_diff %>% group_by(Base) %>% summarise(Count = n())
consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n())

consist_diff %>% filter(Symbol %in% uniq_g$Symbol) %>% group_by(Base) %>% summarise(Count = n()) %>% ggplot(aes(x=`Base`, y = `Count` )) + geom_bar(stat='identity', position = 'dodge') + cowplot::theme_cowplot() + coord_flip() + xlab("CellTye (Predict)")
```

@@

hm <- ComplexHeatmap::Heatmap(x, column_labels = organism, 
                              column_split = cts, 
                              row_split = base_yo, show_row_dend = FALSE, 
                              row_labels = genes, 
                              column_title_rot = 90, row_split
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
ComplexHeatmap::draw(hm, padding = unit(c(2, 2, 32, 2), "mm"), row_title = "Marker Genes")
```

@@

Neurogenic and RPC have very similar expression patterns. We leverage our more specific differential test to directly assess genes were are different between neurogenic cells and RPC. We show the top 5 genes that are both more highly expressed in RPC relative to neurogenic cell and vice versa.
```{r, fig.width=7, fig.height=3}
x <-  scEiaD_2020_v01 %>% 
    tbl("diff_testing") %>% 
    filter(Against  == "RPC",   
           Group == 'CellType (Predict)', 
           Base == "Neurogenic Cell") %>% 
  group_by(Base, Against, Gene) %>% 
  summarise(`mean log2FC` = mean(log2FoldChange), `mean padj` = mean(padj), `sig count` = sum(pvalue < 1e-2), `na count` = sum(is.na(log2FoldChange))) %>% 
  arrange(-`mean log2FC`) %>%   
  collect() %>% 
  filter(!is.na(`mean log2FC`))

@@

@@

hm_flipflop <- ComplexHeatmap::Heatmap(x, column_labels = organism, 
                              row_split = base_yo, show_row_dend = FALSE, 
                              row_labels = genes, 
                              column_title_rot = 90, 
                              name = 'log2FC', use_raster = TRUE,
                              row_title_rot = 0)
hm_flipflop
```
